## Problem statement
In the author column, you will find 74 essays with identified authors: 51 essays written by Hamilton, 15 by Madison, 3 by Hamilton and Madison, 5 by Jay. The remaining 11 essays, however, is authored by “Hamilton or Madison”. These are the famous essays with disputed authorship. In this question, you are provided with the Federalist Paper data set. The features are a set of “function words”, for example, “upon”. The feature value is the percentage of the word occurrence in an essay. 

```{r}
#load packages
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(gridExtra)  # sub-figure layout package
library(dendextend)
library(gridExtra)
```

Load data:
  
  ```{r}
library(readr)
df <- read_csv("~/Desktop/Applied Machine Learning/Week4/HW4-data-fedPapers85(1).csv")
View(df)
```

Analyzing current data

```{r}
#analyze data
#how many documents are from each author
table(df$author)
colnames(df)

```

Search for specific pattern between some of the function words and authros

```{r}
#boxplot of function words: all; upon; into; categorized by author
ggplot(data=df)+aes(y=df$all)+geom_boxplot()+facet_wrap(~author) 
ggplot(data=df)+aes(y=df$upon)+geom_boxplot()+facet_wrap(~author) 
ggplot(data=df)+aes(y=df$into)+geom_boxplot()+facet_wrap(~author) 

```

Pre-processing: remove labeling and normalize dataset and remove all the authors
appart from Hamilton and Madison

```{r}
#remove labeling from data
df_ul<-df[,3:72]

#normalize: mean and standard deviation
df_ul <- scale(df_ul)
df[1:5,1:5]
df_ul[1:5,1:5]

```

```{r}
# Check for missing or non-numeric values
any(is.na(df_ul))
any(!is.numeric(df_ul))



```
```{r}
#nstart = 25 is a value randomly selected
model_4 = kmeans(df_ul, centers = 4, nstart = 25)
#print centroids
model_4$cluster
```


### Unsupervised Learning with K-means
# get concat to the initial dataset the results from clustering
```{r}
cluster_assignment <- data.frame(df,"clusters"=model_4$cluster)
```

### Select the best K

```{r}

k2 <- kmeans(df_ul, centers = 2)
k3 <- kmeans(df_ul, centers = 3)
k4 <- kmeans(df_ul, centers = 4)
k5 <- kmeans(df_ul, centers = 5)


# plots to compare
p2 <- fviz_cluster(k2, geom = "point",  data = df_ul) + ggtitle("k = 2")
p3 <- fviz_cluster(k3, geom = "point",  data = df_ul) + ggtitle("k = 3")
p4 <- fviz_cluster(k4, geom = "point",  data = df_ul) + ggtitle("k = 4")
p5 <- fviz_cluster(k5, geom = "point",  data = df_ul) + ggtitle("k = 5")


grid.arrange(p2, p3, p4, p5, nrow = 2)
```

## Visualizing cluster numbers with authors

```{r}
table(cluster_assignment[,c(1,73)])
```

```{r}
# Subset the cluster_assignment matrix for Hamilton and Madison
ham_mad <- cluster_assignment[cluster_assignment[, 1] %in% c("dispt","Hamilton", "Madison"), ]

# Select the first and 73rd columns
ham_mad_subset <- ham_mad[, c(1, 73)]

# Generate the table of cluster assignments
table(ham_mad_subset)

```

